---
title: "CM_DeerPopModeling_Spring2025"
author: "Conner Mills"
format: html
editor: visual
---

```         
1/15/24

Simple linear models

Components of population change

Nt = Population size at time t

Nt+1 = Population size at the next time step

B = Births

D = Deaths

I = Immigration

E = Emigration

Nt+1 = Nt + B + I - D - E



Assumptions:

Closed population
Nt+1 = Nt + B + I - D - E

Nt+1 = Nt + B - D

$delta$N = Nt+1 - Nt = B - D

Continuous growth
Births and deaths happen continually

$dN/dt$ = B - D

b = instantaneous (per capita) rate = B/N

B = bN

d = D/N

D = dN

$dN/dt$ = B - D = bN - dN = (b-d)N

b and d constant; unlimited resources, no age structure (young and old both have the same birth and death rates), no genetic structure.

Exponential growth model

$dN/dt$ = (b-d)N = rN

r = instantaneous rate of increase (intrinsic rate of increase) - per individual rate of population increase

r > 0?

r = 0

r < 0?

r = $dN/dt$ * 1/N instantaneous contribution each animal makes to population growth

Integrating $dN/dt$ = rN….

Nt = N0 e^rt

N0 initial population size, Nt population size at time t, e is a constant - base of the natural logarithm ~ 2.717

Same equation that banks use to compound interest in your savings account
```

##Quiz Questions## Quiz 1

Estimate the size of a white-tailed deer population for Lee County Alabama in 2030, assuming a population size of 8000 in 2025, r = 0.05.

Write a function in R to calculate population size using the equation Nt = N0 e\^rt .

```{r}
N0=8000
r=0.05
t=5

##Equation##
Nt = N0 * exp(r * t)

##Function##
deerpopFUN = function(N0, r, t) {
  Nt = N0 * exp(r * t)
  return(Nt)
}

deerpop5year = deerpopFUN(N0, r, t)   #error AB
deerpop5year

##Answer = 10,272.2 deer ##
```

Quiz 2

There are 5000 fish in a pond. During a one-month period, you record 750 births and 250 deaths in this population. Assuming population growth is continuous, estimate b, d and r and project the population size in ten months.

```{r}
N0 = 5000 
B  = 750  
D  = 250  
t  = 10

##Equation##
b = B / N0 
b
d = D / N0 
d
r = b - d    
r
Nt = N0 * exp(r * t)

##Function##
fishpopFUN = function(N0, r, t) {
  b = B / N0
  d = D / N0
  r = b - d
  Nt = N0 * exp(r * t)
  return(Nt)
}

fishpop10month = fishpopFUN(N0, r, t)
fishpop10month

##Answers##
##b = 0.15##
##d = 0.05##
##r = 0.1##
##Pop in 10 months = 13591.41##
```
